---
title: "P8105 - Homework 3"
author: "Mingkuan Xu"
output: github_document
---
```{r library}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

#### Part 1 - Data import & description
```{r p1_import}
data("instacart")
head(instacart)
```

This is an anonymized dataset recording online grocery orders from Instacart users. Each row in the dataset is a product from an order; there is a single order per user in this dataset. There are `r nrow(instacart)` total rows in this dataset with `r ncol(instacart)` columns, recording `r length(unique(pull(instacart,order_id)))` orders in total. The recorded information include: `r tolower(names(janitor::clean_names(instacart,"sentence")))`. There are some interesting observations of this dataset. For example, there are `r length(unique(pull(instacart,department)))` departments in total, including `r unique(pull(instacart,department))`.

#### Part 2 - Analysis of aisles

```{r p1_part2}
instacart %>% pull(aisle) %>% unique() %>% length() # Find number of aisles.

instacart_by_aisle = 
  instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) 

head(instacart_by_aisle,1)# Find most popular aisle
```

Using the codes above, we found that there are 134 aisles in total, with aisle of the fresh vegetables being the most popular one from which 150609 items were ordered.

#### Part 3 - Plot of items from aisles.

Now we plot the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered using the following code.

```{r p1_part3}
instacart_by_aisle %>% 
  filter(n>10000)  %>%
  ggplot(.,aes(x=reorder(aisle,n),y=n)) + 
  geom_bar(stat="identity") +
  xlab("Aisle")+
  ylab("Number of Items") +
  coord_flip()
```

#### Part 4 - Three most popular items

```{r p1_part4}
baking_ingredients = 
  instacart %>%
  filter(aisle=="baking ingredients") %>%
  count(product_name) %>%
  arrange(desc(n)) %>%
  mutate(Aisle="Baking Ingredients") %>%
  rename(Product=product_name,Count=n)%>%
  relocate(Aisle) %>%
  head(3)
  

dog_food_care = 
  instacart %>%
  filter(aisle=="dog food care") %>%
  count(product_name) %>%
  arrange(desc(n)) %>%
  mutate(Aisle="Dog Food Care") %>%
  rename(Product=product_name,Count=n)%>%
  relocate(Aisle) %>%
  head(3)

packaged_vegetables_fruits = 
  instacart %>%
  filter(aisle=="packaged vegetables fruits") %>%
  count(product_name) %>%
  arrange(desc(n)) %>%
  mutate(Aisle="Packaged Vegetables Fruits") %>%
  rename(Product=product_name,Count=n)%>%
  relocate(Aisle) %>%
  head(3)

popular_table = 
  bind_rows(baking_ingredients,
                          dog_food_care,
                          packaged_vegetables_fruits) %>%
  arrange(desc(Count))

popular_table %>% knitr::kable(digits = 1)
```

#### Part 5 - The mean hour of the day

```{r p1_part5}

pink_lady_apples = 
  instacart %>%
  filter(product_name == "Pink Lady Apples") %>%
  select(order_dow,order_hour_of_day) %>%
  group_by(order_dow) %>%
  summarize(
    median_hour_of_day = mean(order_hour_of_day)
  ) %>%
  rename("Pink Lady Apples"=median_hour_of_day)


coffee_ice_cream = 
  instacart %>%
  filter(product_name == "Coffee Ice Cream") %>%
  select(order_dow,order_hour_of_day) %>%
  group_by(order_dow) %>%
  summarize(
    median_hour_of_day = mean(order_hour_of_day)
  ) %>%
  rename("Coffee Ice Cream"=median_hour_of_day)

median_hour_table = 
  left_join(pink_lady_apples,coffee_ice_cream,by="order_dow") %>%
  select(-order_dow)
# mutate("Day of Week" = c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

median_hour_table %>% knitr::kable(digits = 1)
```

## Problem 2
```{r p2_import}
data("brfss_smart2010")

data_brfss = brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic=="Overall Health") %>%
  filter(response %in% c("Excellent","Very good","Good","Fair","Poor"))

nrow(data_brfss)
# unique(pull(data_brfss,response))
```

